---
title: "United Nations Voting Records"
author: "River Mckegney"
format: 
  html: default
  gfm: default
embed-resources: true
code-tools: true
toc: true
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r}
#| label: load packages
#| include: false
library(tidyverse)
library(lubridate) #should already be installed as a part of the tidyverse, we will talk more about this in Week 10
library(scales)  #you may need to install package `scales`
```


***This task is complex. It requires many different types of abilities. Everyone will be good at some of these abilities but nobody will be good at all of them. In order to produce the best product possible, you will need to use the skills of each member of your group.***

<!-- The person who has the most siblings (or cousins if both are only children) starts as the Developer (typing and listening to instructions from the Coder)!  -->


::: {.callout-note}
Note: Since the project folder is already shared as a zip file, using the commands:  

  - `usethis::use_git()`  
  - `usethis::use_github()`

might be easier to initiate your github repository. 
:::


## Goals for the Activity  

- Join multiple data tables together by a common variable(s)  
- Create new data sets through the joining of data from various sources  
- Combine `join` functions with other `tidyverse` functions  

**THROUGHOUT THE Activity** be sure to follow the Style Guide by doing the following:  

- load the appropriate packages at the beginning of the Quarto document  
- use proper spacing  
- *add labels* to all code chunks  
- comment at least once in each code chunk to describe why you made your coding decisions  
- add appropriate labels to all graphic axes  




# Data Description
The data this week comes from Harvard's Dataverse by way of Mine Çetinkaya-Rundel, David Robinson, and Nicholas Goguen-Compagnoni.

> Original Data citation: Erik Voeten "Data and Analyses of Voting in the UN General Assembly" Routledge Handbook of International Organization, edited by Bob Reinalda (published May 27, 2013). Available at SSRN: http://ssrn.com/abstract=2111149

It was featured on [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-23/readme.md)

Here is each data set and its description (you might want to look at the Tidy Tuesday link for the tables already rendered)

`unvotes.csv`

|variable     |class     |description |
|:------------|:---------|:-----------|
|rcid         |double    | The roll call id; used to join with un_votes and un_roll_call_issues |
|country      |character | Country name, by official English short name |
|country_code |character | 2-character ISO country code |
|vote         |integer   | Vote result as a factor of yes/abstain/no |


```{r}
#| label: Read in dataset unvotes
#| echo: true
#| output: false

# Call dataset unvotes as a tibble
unvotes <- read_csv("data/unvotes.csv") 
```



`roll_calls.csv`

|variable      |class     |description |
|:-------------|:---------|:-----------|
|rcid          |integer   |.           |
|session       |double    | Session number. The UN holds one session per year; these started in 1946|
|importantvote |integer   | Whether the vote was classified as important by the U.S. State Department report "Voting Practices in the United Nations". These classifications began with session 39|
|date          |double    | Date of the vote, as a Date vector|
|unres         |character | Resolution code |
|amend         |integer   | Whether the vote was on an amendment; coded only until 1985 |
|para          |integer   | Whether the vote was only on a paragraph and not a resolution; coded only until 1985|
|short         |character |  Short description |
|descr         |character | Longer description|


```{r}
#| label: Read in dataset roll_calls
#| echo: true
#| output: false

# Call dataset roll_calls as a tibble
roll_calls <- read_csv("data/roll_calls.csv") 
```



`issues.csv`

|variable   |class     |description |
|:----------|:---------|:-----------|
|rcid       |integer   | The roll call id; used to join with unvotes and un_roll_calls |
|short_name |character | Two-letter issue codes |
|issue      |integer   | Descriptive issue name |

```{r}
#| label: Read in dataset issues
#| echo: true
#| output: false

# Call dataset issues as a tibble
issues <- read_csv("data/issues.csv")
```



<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->



## Data Exploration
Our goal today is to explore how various members (countries) in the United Nations vote. We have three data sets, what can we determine from each data set separately?

Response by River Mckegney
> With the 'unvotes' dataset we can evaluate location data (country level) associated with votes and also a countries vote type.
With the 'roll_calls' dataset we can suss out the datetime and information about vote meetings held by the United Nations. The 'issues' dataset contains pertinent information about the vote topic.

### UN Votes

The first data set, `unvotes` contains data on the `rcid` which is the roll call id for the vote, the country/country code, and how the country voted. What can we learn from the data?  

Response by River Mckegney
> We can learn how each country decidided to vote about a specific issue, and what similarities or differences exist between countries representitive views.


Comment on the following code, what is happening in each line?  One way to approach seeing what each line does is to highlight the code from before the `pipe` of that line up to the data `unvotes` and use `CTRL + ENTER` to run just the highlighted lines.  

```{r}
#| label: unvotes barplot
#| echo: true
#| output: true

unvotes |> 
  count(country, vote) |> # counts total vote types for each country
  group_by(country) |> # groups next calculation by country
  mutate(total = sum(n)) |> # Counts all votes from all types for each  country
  mutate(prop_vote = n/total) |> # Calculates proportion of each vote type
  filter(country %in% c("United States", "Canada", 
                        "Germany", "France",
                        "Italy", "Japan",
                        "United Kingdom")) |> # Selects variables from these specific countries
  ggplot(aes(x = country, y = prop_vote, 
             fill = vote)) + # Creates a plot with x-axis as countries and y-axis as proportions of votes
  geom_col(position = position_stack()) + # Creates a stacked bar chart
  labs(x = "Group of Seven Countries",
       y = "Proportion of Votes",
       title = "Voting Record of the G7",
       fill = "Vote") + # Labels each axis, and color codes bars by vote type
  theme_minimal() +  # Adds a minimal theme to graph
  scale_fill_viridis_d(end = 0.8) + # Adds specific color palette to chart
  coord_flip() # Flips the axes and stacked bar plot orientation
```

Describe what the graph above demonstrates above UN voting records for the G7 

Response by River Mckegney
>  From the barplot we can see that from these seven selected couintries, looking at the historical extent of all UN voting that happend (in the dataset), The United States contains the highest count for voting "no" and Germany has the highest count for voting "yes". 
  
  

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->




### Roll Calls

The second data set, `roll_calls` has more information on the type of vote, the importance, whether it was a resolution, and date of the vote.  What does the following code do?

```{r}
#| label: Unique values of short col in roll_calls
#| echo: true
#| output: true

roll_calls |> 
  distinct(short)
```

Description of code results

Response by River Mckegney
> The code block above displays the unique variables (short response descriptions) for the voting topics being decided upon. 



We can use the individual data for `roll_calls` to look at the number of votes per year over time.

```{r}
#| label: 
#| echo: true
#| output: true

roll_calls |> 
  mutate(year = lubridate::year(date)) |> #extracts the year from the date value and creates a new `year` column
  count(year) |> #counts how many votes there were per year assuming each line is an single voting instance
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  labs(x = "Year", y = "Number of Votes",
       title = "UN Votes per Year") +
  theme_minimal()
```



What information is missing from the above graphic that might be useful in understanding the issues the UN commonly votes on?

Response by River Mckegney
> The graph is missing information about the voting topics, most appropriately the short descriptions associated with each UN ballot event in the 'short' column of the roll_calls dataset.



We could try and look at the short descriptions, `short`, for each vote with the following code. 
```{r}
roll_calls |> 
  count(short)
```

Does the above information help us understand the voting issues of the UN over time? Explain.

Response by River Mckegney
> Yes, but not quite. The code above counts the total number of discrete vote topics, but does not divy out these counts across the years that they occured. This filtering process could be easily used with the graph above (and perhaps another grouping or filtering function) to illustrate how UN voting counts on specific issues change over time.



<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->


### Issues 
Finally we have the `issues` data which provides a more general description for each vote on specific issues.  Note that not all issues are included in the data set, just the ones related to the 6 issues below:

```{r}
issues |> 
  count(issue)
```

It would be helpful to use the `issues` data with the `roll_calls` data to be able to better understand the voting trends within the UN on at least these 6 issues. To do this, we need to join the data.  



# Votes Over Time  
Now let's join our data together to get a better idea of how the UN has voted over time.  
First, look at the number of rows in `issues` and `roll_calls` - do they match? What does this indicate?

```{r}
dim(issues)
dim(roll_calls)
```

Response by River Mckegney
> The number of rows do not match between 'issues' (5745 rows) and 'roll_calls' (6202 rows). This indicates that the issues dataset does not contain all the observations (the historical record of votes data) that roll_calls does.



Now let's try joining the `roll_calls` with the `issues` data. Compare the following codes to join the data. Describe what each one does and how it differs from the others as a comment in the code chunk.  You might need to reference the slides or reading for this week.

```{r}
roll_calls |> 
  left_join(issues, by = "rcid") 

# Left joins merge the second or right df y 'issues' by all rows matching the first left df x 'roll_calls'. Missing values of df y across rows are then filled in with NAs.

# Left join = keep all x rows, remove non-matching y rows
```

```{r}
roll_calls |> 
  right_join(issues, by = "rcid") 

# Right joins will keep all rows of df y 'issues' and remove unmatched rows of df x 'roll_calls' 

# Right join = keep all y rows, remove unmatched x rows
```

```{r}
roll_calls |> 
  full_join(issues, by = "rcid") 

# Full joins will display all rows of left df x 'roll_calls' first, as is, then will include any unmatched rows of right df y 'issues' 

# Full join = keep both x and y rows, including unmatched ones
```

```{r}
roll_calls |> 
  inner_join(issues, by = "rcid") 

# Inner join = Keep only matching rows in both x and y
```

**If we are only interested in retaining the records associated with the `issues` labeled in our data, which join(s) should we use?**

Response by River Mckegney
> With this format (assuming roll_calls as left df, x), we could use a right join to retain all rows in issues (removes unmatched rows from roll_calls), or a full join to keep rows from both dataframes. 




<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->


Now that we know how to join the data, we will use the following code to examine the the voting trends for three of the issues related to conflict/weapons. 

Be sure to run the code via the green arrow on the code chunk, as the `case_when()` code can get finicky sometimes and claim an error about a comma in the code when it doesn't exist.  *Comment on the code where indicated* and add your chosen join function
```{r}
roll_calls |> 
  right_join(issues, by = "rcid") |>  #join roll_calls and issues so that just the votes related to the issues data are retained.
  mutate(issue = case_when(
    issue == "Arms control and disarmament" ~ "Arms Control",
    issue == "Nuclear weapons and nuclear material" ~ "Nuclear Weapons",
    issue == "Palestinian conflict" ~ "Palestinian Conflict")) |>   # Here we select and rename each of the variables or topics of interest within the issue column, which effectively bins or groups subjects into a neater format for graphic display
  
  drop_na(issue) |> #drop NAs for values not defined in the previous line of code
  mutate(year = lubridate::year(date)) |> #create a column `year` that contains the year value
  count(year, issue) |> # Counts the number of votes of a particular issue for a given year
  ggplot(aes(x = year, y = n, group = issue)) + # Creates a graph and id's axes
  geom_line(aes(color = issue)) + # Makes line chart and color codes by issue variables
  labs(x = "Year", y = "Number of Votes",
       title = "United Nations Votes per Year",
       subtitle = "Conflict and Arms Related Votes",
       color = "Voting Issue") +
  theme_minimal()
```




What do you notice? What do you wonder based on the graph created?

Response by River Mckegney
> I notice that the trendlines for these debated topics tend to correlate or set eachother off. So for spike around 1985 in arms controls there is also lots of voting going on about the palestian conflict and nuclear weapons. Based on the graph I also notice nuclear weapons is the topic that generally tends to display the highest voting counts. And, on average voting has increased with time. 




<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->


# Challenge 1: Joining all Data

We want to try and create the visualization (see instructions online).  To do this, though, we need information from all three data set, `unvotes`, `issues`, and `roll_calls`

We want to join all three data sets together, *maintaining only the votes for which we have identified the general issue* (e.g., Nuclear War, Arms, Economics, etc.).  We will save (assign) the data as `un_full`.  Fill in the best *join functions* to do maintain just the votes for which we have identified the general issue.

```{r}
un_full <- unvotes |> 
  right_join(issues, by = "rcid") |> # Keep 'issues' rows, remove unmatched unvotes rows
  left_join(roll_calls, by = "rcid") # Retain left x rows, remove unmatched y rows

#View(un_full)
```


Now, we are going to do some data cleaning.  Our goal is create a data set that includes the percentage of "yes" votes per country each year. Fill in the missing _________ with the correct verbs. We will call the data table `yes_votes`
```{r}
yes_votes <- un_full |> 
  select(country, issue, date, vote) |>  #keep just the columns for country, issue, date, vote
  mutate(year = lubridate::year(date)) |> #create a new variable called year
  group_by(country, year, issue) |> #group country, year, issue together to prepare for subsequent analysis
  summarize(percent_yes = mean(vote == "yes")) |> #calculate the proportion of yes votes
  mutate(issue = case_when(
    issue == "Arms control and disarmament" ~ "Arms Control",
    issue == "Nuclear weapons and nuclear material" ~ "Nuclear Weapons",
    issue == "Palestinian conflict" ~ "Palestinian Conflict",
    TRUE ~ issue)) #create/overwrite issue column with simpler values

#View(yes_votes) #you can use this code to look at the table created
```


Now write the clean data `yes_votes` to a new data set in the data folder so we don't have to rerun the above code again if we want to use the data in the future.  


```{r}
write_csv(yes_votes, "data/yes_votes.csv")
```




Now we can feed the `yes_votes` transformed data table into your graphing code, but first we will want to focus on the United States and Canada. *Provide a comment to describe what each line of code is doing in the process*.  

```{r}
yes_votes |> 
  filter(country %in% c("United States","Canada")) |> # select only countries from US and CA
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) + # Create graph, set x & y axis, color code by country
  geom_point(alpha = 0.4) + # Create scatterplot, set transparency 
  #geom_line(aes(group = country)) +
  geom_smooth(method = "loess", se = FALSE) + #this fits a special model called a loess regression, a smooth line that fits the data
  facet_wrap(~issue) + # Divy out plots based on issue variable
  scale_y_continuous(labels = scales::percent) + # Change y-axis scale to percent values
  labs(
    title = "Percentage of 'Yes' votes in the UN General Assembly",
    subtitle = "1946 to 2015",
    y = "% Yes",
    x = "Year",
    color = "Country"
  ) + # Edit and add labels for graph title, subtitle, x & y axis, and legend title
  theme_bw() +
  theme(legend.position = "bottom") + # Move legend position to bottom of graph
  scale_color_viridis_d(option = "turbo") # Adjust the color scheme 
```


What do you notice about the voting records over time? 

Response by River Mckegney
> Trends across all voting topics tend to follow a similar rise and fall pattern (of yes votes) which repreats itself through time. Topics of arms control, human rights, and nuclear weapons appear to be correlated as they follow similar trends relative to the timescale. Likewise, colonialism, economic development, and the palestinian conflict also appear related to eachother.  






# Challenge 2: Adding more data
<!-- You do not need to complete this section unless you want to! It is a challenge problem to help you further apply what you did above. -->
After your instructor created the above plot, she became curious about how politics might impact the UN Voting record for the United States since UN Ambassador is a presidential appointment. So your instructor started searching for a data set of US presidents, their years in office, and their political affiliation.  She found a data set on Kaggle.com and removed the information prior to 1940 (because the dates were coded funny and it was causing problems). She saved the data as `us_presidents.csv` and imported it into the project.

```{r}
president <- read_csv("data/us_presidents.csv")
```

She realized that her data only had the start/end dates for each president and she wanted a data set that filled in the missing years and political party for the president in that time period.  After much googling and reading Stack Overflow, found two functions she did not know about called `complete()` and `fill()` to fill in the missing years and party affiliations  


```{r}
politics_year <- president |> 
    mutate(start = lubridate::mdy(start),  #formats date correctly
           start_year = lubridate::year(start)) |> #pulls out year
  filter(start_year > 1940) |>  #removes data before 1940 since there was no UN
  select(start_year, party) |>   #pulls out just the variables of interest
  complete(start_year = seq(min(start_year), 2020, by = 1)) |> #fill in missing years
  fill(party) #fill in missing party affiliations for years
```


Next, your instructor took the `yes_votes` data and filtered out just the United States data and then joined by year to add the party affiliation of the president for each year of UN votes.  To create the visualization with the smoothed model, but color by party affiliation, she had to add a new column called `predict` that fit the model first instead of using `geom_smooth()`.  

Comment the code below: 
```{r}
yes_votes |> 
  filter(country == "United States") |>  #comment
  left_join(politics_year, by = c("year" = "start_year")) |> #comment
  group_by(issue) |>  #comment
  mutate(predict = predict(loess(percent_yes~year))) |>  #comment
  ggplot(mapping = aes(x = year, color = party, group = 1)) + #comment
  geom_point(aes(y = percent_yes),alpha = 0.4) + #comment
  geom_line(aes(y = predict)) + #comment
  facet_wrap(~issue) + #comment
  scale_y_continuous(labels = scales::percent) + #comment
  labs(
    title = "Percentage of 'Yes' votes in the UN General Assembly",
    subtitle = "1946 to 2015",
    y = "% Yes",
    x = "Year",
    color = "Country"
  ) + #comment
  theme_bw() + #comment
  scale_color_manual(values = c("blue", "red")) + #comment
  theme(legend.position = "bottom")  #comment
```

What do you notice? What do you wonder?

> Insert your answer


